---
title: "Practice: Taxonomy of Data"
---

## 1) Types of Variables

For each of the following parts indicate the type of variable according to the taxonomy of data.

a) The number of classes a student is taking

b) A student's year in college (Freshman, Sophomore, etc.)

c) Student's favorite color

d) The area code of a student’s phone number

e) The amount of time it takes to finish a STAT 20 quiz

f) The temperature of a room in Fahrenheit degrees

g) The type of car you drive

h) The speed of a vehicle in miles per hour

i) A car's model year

j) The number of text messages you send per day

k) The brand of coffee you drink

l) Your student ID number

m) The number of siblings a person has

n) A person's home zip code

o) The rating of a movie (1 to 5 stars)

::: {.callout-note collapse="true"}
## Show answers

a) discrete numerical
b) ordinal categorical
c) nominal categorical
d) nominal categorical
e) continuous numerical
f) continuous numerical
g) nominal categorical
h) continuous numerical
i) ordinal categorical
j) discrete numerical
k) nominal categorical
l) nominal categorical
m) discrete numerical
n) nominal categorical
o) usually discrete numerical, but one can also argue that it can be ordinal categorical (depending on the context, and application)
:::


## 2) Terrestrial Planets

Consider the following table containing data from so-called **Terrestrial**
planets.

| name    | gravity | temp  | moons_num | moons_fac | haswater |
|:-------:|:-------:|:-----:|:---------:|:---------:|:--------:|
| Mercury |  3.7    | 167   |   0       | none      |  no   |
| Venus   |  8.9    | 464   |   0       | none      |  no   |
| Earth   |  9.8    |  15   |   1       |  one      |  yes  |
| Mars    |  3.7    | -65   |   2       |  two      |  no   |

- `name`: name of planet
- `gravity`: gravity ($m/s^2$)
- `temp`: mean temperature in Celsius
- `moons_num`: number of moons (numeric)
- `moons_fac`: number of moons (factor)
- `haswater`: whether it has known bodies of liquid water on its surface


### 2.1) Creating Vectors and Factors

Go to `stat20.datahub` and open a new Quarto document (`.qmd` file).

Refer to the data table of terrestrial planets. Insert new code chunks to 
create the following vectors.

a) Create a vector `name` and display it (i.e. type its name)

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
name <- c("Mercury", "Venus", "Earth", "Mars")
name
```

b) Create a vector `gravity`

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
gravity <- c(3.7, 8.9, 9.8, 3.7)
gravity
```


c) Create a vector `temp`

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
temp <- c(167, 464, 15, -65)
temp
```


d) Create a vector `moons_num`

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
moons_num <- c(0, 0, 1, 2)
moons_num
```


e) Create a factor `moons_fac`

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
moons_fac <- factor(c("none", "none", "one", "two"))
moons_fac
```


f) Create a factor `haswater`

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
haswater <- factor(c("no", "no", "yes", "no"))
haswater
```


### 2.2) Some basic computations

Answer each of the following parts in their own code chunks.

a) Use the `length()` function to find the number of elements in vector `name`.

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
length(name)
```


b) Use the `sqrt()` function to calculate the square root of gravity values.

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
sqrt(gravity)
```


c) Use the `mean()` function to compute the mean number of moons.

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
mean(moons_num)
```


d) Use the comparison operator `>` to see which temperature values in `temp` 
are greater than 100

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
temp > 100
```


e) Use the following formula to compute `x`

$$
\texttt{x} = \frac{\texttt{temp}^2 + \log(\texttt{gravity})}{\texttt{moons\_num} + 10}
$$

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
x = (temp^2 + log(gravity)) / (moons_num + 10)
x
```

\

### 2.3) Data Frame `planets`

In a new code chunk, use the `data.frame()` function to create a `planets` 
data frame containing all the vectors and factors created in part 2.1.
And then type (print) `planets` in order to see its contents. 

```{r}
#| code-fold: true
#| code-summary: "Show answer"
#| eval: false
planets = data.frame(name, gravity, temp, moons_num, moons_fac, haswater)
planets
```

