---
title: "Taxonomy of Data"
format: stat20slides-revealjs
chalkboard: true
---

## Agenda

-   Announcements
-   Group Activity: Conceptual
-   Worksheet 1: *Taxonomy of Data*
-   *Break*
-   Concept Questions: Coding
-   More Practice

## Assignments 

RQ: Taxonomy of Data (due yesterday Thu 1/22 11:59pm)

. . .

\

Quiz-1 (Monday 1/26)

. . .

\

Lab-1 **Getting Started** (due Tuesday 1/27 @7am)

. . .

\

Problem Sets (due Tuesday 1/27 @7am)

- PS0: 2 google forms
- PS1: _Understanding the World with Data_ + _Taxonomy of Data_


## Lab 1: Getting Started

- Complete `qmd` file

- Render `qmd` to obtain PDF document

- Download (_export_) `pdf` to your computer (or tablet)

- Submission: upload PDF to gradescope


# Notes Recap

## Concept of Data

One or more **characteristics** observed or measured on a set of **objects**.

## Concept of Data

One or more ~~characteristics~~ observed or measured on a set of objects.

. . .

-   Variables: characteristics, features, attributes.

-   Objects: individuals, subjects, items.

\

. . .

Data typically organized into a data table (or *data frame*), ideally one row per individual, and one column per variable.

## Types of Variables

Variables can be classified in different ways.

. . .

The classification (*aka* taxonomy) adopted in STAT 20 involves 2 major classes, each one with two subcategories:

. . .

:::: {.columns}
:::{.incremental}
::: {.column width="50%"}
-   Numerical

    -   Continuous
    -   Discrete
:::

::: {.column width="50%"}
-   Categorical

    -   Ordinal
    -   Nominal
:::
:::
::::


# Group Activity

##  {data-menu-title="Variables"}

1.  As a group, choose a set of individuals, and list at least:

    -   3 continuous numerical variables
    -   3 discrete numerical variables
    -   3 ordinal categorical variables
    -   3 nominal categorical variables.

2.  ❌ Do Not choose:

    -   students
    -   electronic devices

```{r}
countdown::countdown(15, top = 0)
```


# Worksheet 1: *Taxonomy of Data*

<https://stat20.berkeley.edu/spring-2026/1-questions-and-data/02-taxonomy-of-data/ws.pdf>

```{r}
countdown::countdown(20, bottom = 0)
```

# *Break*

```{r}
countdown::countdown(5, top = 0)
```


# Practice Problems

[www.gastonsanchez.com/stat20/1-questions-and-data/02-taxonomy-of-data/practice.html](https://www.gastonsanchez.com/stat20/1-questions-and-data/02-taxonomy-of-data/practice.html)

These are NOT worksheets (no need to submit to Gradescope)

```{r}
countdown::countdown(25, top = 0)
```

