---
title: "Taxonomy of Data"
format: stat20slides-revealjs
chalkboard: true
---

## Agenda

-   Announcements
-   Group Activity: Conceptual
-   Worksheet 1: *Taxonomy of Data*
-   *Break*
-   Concept Questions: Coding
-   More Practice

## Assignments

You should have submitted Lab-1 **UN Votes** and Worksheet Packet WSP-1

These were originally due yesterday (Tue) on Gradescope by 8:00am, but got extended till 11:59pm.

. . .

\

You should have answered Reading Questions of **Taxonomy of Data**

These were due yesterday (Tue) on Gradescope by 11:59pm

# Notes Recap

## Concept of Data

One or more **characteristics** observed or measured on a set of **objects**.

## Concept of Data

One or more ~~characteristics~~ observed or measured on a set of objects.

. . .

-   Variables: characteristics, features, attributes.

-   Objects: individuals, subjects, items.

\

. . .

Data typically organized into a data table (or *data frame*), ideally one row per individual, and one column per variable.

## Types of Variables

Variables can be classified in different ways.

. . .

The classification (*aka* taxonomy) adopted in STAT 20 involves 2 major classes, each one with two subcategories:

. . .

:::: {.columns}
:::{.incremental}
::: {.column width="50%"}
-   Numerical

    -   Continuous
    -   Discrete
:::

::: {.column width="50%"}
-   Categorical

    -   Ordinal
    -   Nominal
:::
:::
::::


# Group Activity

##  {data-menu-title="Variables"}

1.  As a group, choose a set of individuals, and list at least:

    -   3 continuous numerical variables
    -   3 discrete numerical variables
    -   3 ordinal categorical variables
    -   3 nominal categorical variables. Don't choose:

2.  ❌ Do Not choose:

    -   students
    -   electronic devices

```{r}
countdown::countdown(15, top = 0)
```


# Worksheet 1: *Taxonomy of Data*

<https://stat20.berkeley.edu/fall-2025/assignments.html>

```{r}
countdown::countdown(20, bottom = 0)
```

# *Break*

```{r}
countdown::countdown(5, top = 0)
```


# Practice Problems

Additional handout prepared for Prof. Sanchez's sections (1 & 8)

These are NOT worksheets (no need to submit to Gradescope)

```{r}
countdown::countdown(25, top = 0)
```

# Your Turn

::: poll
1.  Create a vector named `vec` with the even integers between 1 and 10 as well as the number 99 (six elements total).

2.  Find the sum of that vector.

3.  Find the max of that vector.

4.  Take the mean of that vector and round it to the nearest integer.
:::

These should all be solved with R code. If you don't know the name of a function to use, you could hazard a guess by looking for a help file (e.g. ?sum) or google it.

```{r}
countdown::countdown(5, top = 0)
```

# Your Turn

::: poll
1.  Create a new .qmd file, name it, and save it.

2.  Insert a new code cell.

3.  Create three vectors, `name`, `hometown`, and `sibs_and_pets` that contain observations on those variables from 6 people in this class.

Combine them into a data frame called `my_classmates.`
:::

```{r}
countdown::countdown(6, top = 0)
```

# Reminder

Answer Reading Questions of **Summarizing Categorical Data**

Due on Gradescope by 11:59pm Thursday
