---
title: "Summarizing Numerical Data"
format: stat20slides-revealjs
chalkboard: true
---

## Agenda

-   Announcements
-   Notes Recap
-   Practice Concept Questions
-   Coding Activity: Graphing Numerical Data
-   *Break*
-   Worksheet "Summarizing Numerical Data" (part of WSP-3)
-   Lab 2.2 (*time permitting*)

## Announcements

-   RQ: Grammar of Graphics due **Thursday at 11:59pm**

. . .

-   Lab 2: Class Survey (both parts) due **Tuesday at 8am**

    -   Lab 2.1: Group submission
    -   Lab 2.2: Individual submission


# Notes Recap

Summarizing Distributions of Data using:

-   Graphics
-   Numerical Summaries

## Graphics

You can construct a **statistical graphic** to show the [shape]{.hilit}, which you can describe in terms of [modality]{.bold-hilit} and [skew]{.bold-hilit}

-   Dot plot
-   Histogram
-   Density plot
-   Violin plot
-   Box plot


## Measures of Center

You can calculate a [measure of center]{.bold-hilit} to convey a sense of a typical (representative) observation

-   Mean
-   Median
-   Mode


## Measures of Spread

And you can calculate a [measure of spread]{.bold-hilit} (i.e. scatter, dispersion, variation) to capture how much variability there is in the data

-   Range
-   Inter Quartile Range (IQR)
-   Mean Absolute Deviation (MAD)
-   Sample Variance (Var)
-   Sample Standard Deviation (SD)


## 

Typical value?

$$ 6 \quad 7 \quad 7 \quad  7 \quad 8 \quad  8 \quad 9 \quad 9 \quad 10 \quad 11  \quad 11$$

. . .

-   sample size ( $n$ ): 11
-   mean ( $\bar{x}$ ): 8.45
-   median: 8
-   mode: 7


## 

How can we express the variability in this data set using a single number?

$$ 6 \quad 7 \quad 7 \quad  7 \quad 8 \quad  8 \quad 9 \quad 9 \quad 10 \quad 11  \quad 11$$

## 

$$ {\Large 6} \quad 7 \quad 7 \quad  7 \quad 8 \quad  8 \quad 9 \quad 9 \quad 10 \quad 11  \quad {\Large 11}$$

. . .

### The Range

$$\textrm{range:} \quad max - min$$

$$ 11 - 6 = 5$$

**Characteristics**

-   Very sensitive to extreme values!


## 

$$ 6 \quad 7 \quad {\Large 7 \quad  7} \quad 8 \quad  {\large 8} \quad 9 \quad {\Large 9 \quad 10} \quad 11  \quad 11$$

. . .

### The Inner Quartile Range (IQR)

The difference between the 3rd quartile, $Q_3$, and the 1st quartile, $Q_1$ (i.e. the middle 50% of the data)

$$\textrm{IQR:} \quad Q_3 - Q_1$$

$$ 9.5 - 7 = 2.5 $$

**Characteristics**

-   Robust to outliers
-   Used to set the width of the box in a boxplot

## 

$$ 6 \quad 7 \quad 7 \quad  7 \quad 8 \quad  8 \quad 9 \quad 9 \quad 10 \quad 11  \quad 11$$

. . .

### Mean Absolute Deviation

Take the differences from each observation, $x_i$, to the sample mean, $\bar{x}$, take their absolute values, add them up, and divide by $n$. Simply put, this is the average distance from the mean.

$$MAD: \quad \frac{1}{n}\sum_{i = 1}^n |x_i - \bar{x}| $$

```{r eval = FALSE}
x <- c(6, 7, 7, 7, 8, 8, 9, 9, 10, 11, 11)
sum(abs(x - mean(x)))/11
```

$$ MAD = 1.4 $$

**Characteristics**

-   Incorporates information from all observations
-   Robust to extreme values

## 

$$ 6 \quad 7 \quad 7 \quad  7 \quad 8 \quad  8 \quad 9 \quad 9 \quad 10 \quad 11  \quad 11$$

. . .

### Sample Variance

Take the differences from each observation, $x_i$, to the sample mean, $\bar{x}$, square them, add them up, and divide by $n - 1$ .

$$s^2: \quad \frac{1}{n - 1}\sum_{i = 1}^n (x_i - \bar{x})^2 $$

```{r eval = FALSE}
x <- c(6, 7, 7, 7, 8, 8, 9, 9, 10, 11, 11)
sum((x - mean(x))^2)/(11-1)
var(x)
```

$$ s^2 = 2.87 $$

:::::: columns
**Characteristics**

::: {.column width="33%"}
-   Incorporates information from all observations
:::

::: {.column width="33%"}
-   Moderately sensitive to extreme values
:::

::: {.column width="33%"}
-   Measured in squared units!
:::
::::::

## 

$$ 6 \quad 7 \quad 7 \quad  7 \quad 8 \quad  8 \quad 9 \quad 9 \quad 10 \quad 11  \quad 11$$

. . .

### Sample Standard Deviation

Take the differences from each observation, $x_i$, to the sample mean, $\bar{x}$, square them, add them up, divide by $n - 1$, then take the square root.

$$
S:  \sqrt{\frac{1}{n - 1}\sum_{i = 1}^n (x_i - \bar{x})^2}
$$

```{r eval = FALSE}
x <- c(6, 7, 7, 7, 8, 8, 9, 9, 10, 11, 11)
sqrt(sum((x - mean(x))^2)/(11-1))
sd(x)
```

$$ s = 1.70 $$

:::::: columns
**Characteristics**

::: {.column width="33%"}
-   Incorporates info from all observations
:::

::: {.column width="33%"}
-   Moderately sensitive to extreme values
:::

::: {.column width="33%"}
-   Measured in units of the original data
:::
::::::

# Practice Concept Questions

```{r}
countdown::countdown(15, top = 0)
```

# Introducing `ggplot2`

Demo

# tinyurl.com/ybhwtrr9

Coding Activity: Graphing Numerical Data

```{r}
countdown::countdown(minutes = 25, top = 0)
```

# Break

```{r}
countdown::countdown(minutes = 5, top = 0)
```

# Worksheet: Summarizing Numerical Data

```{r}
countdown::countdown(minutes = 30, top = 1)
```

# Lab 2.2 Time Permitting

# End of Lecture
