---
title: "Practice: Normal Approximation"
---


```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## 1) Rolling Dice

A fair die will be rolled $n$ times and the object is to guess the total number 
of spots in $n$ rolls. For which value of $n$ will the SE be greater: 

a) $n = 100$

b) $n = 200$

c) SE will be the same regardless of $n$


## 2) Random Variables

We have a box of 50 tickets marked with one of {-2, 1, 0, 1, 2}. We draw $n$ 
tickets at random with replacement from this box. Which of the following 
quantities is random? Select all that apply.

a) The average of the tickets in the box

b) The variance of the tickets in the box

c) The average $\bar{X}$ of the draws. 

d) The sum $S_n$ of the draws


::: {.callout-note collapse="true"}
## Show answers

1) option a) will give you smaller SE
2) options c) and d)
:::


## 3) CLT

Consider a `box` with 50 tickets of which 40 are marked with "0" and 10 are marked with "1". 
Suppose we draw 100 tickets---with replacement---from the `box` and look at the 
sum of the draws $Sn$; like in the following code:

```{r}
set.seed(468)
box <- c(rep(0, 40), rep(1, 10))

draws = sample(box, size = 100, replace = TRUE)
sum(draws)
```


Write R code, using the function `replicate()`, to generate 5,000 simulations of the above chance process, and make a histogram of the 5,000 sums. Does the Central Limit Theorem applies in this case?

::: {.callout-note collapse="true"}
## Show answers

```{r}
sums = replicate(
  n = 5000,
  expr = {
    draws = sample(box, size = 100, replace = TRUE)
    sum(draws)
  }
)

data.frame(sums) |> 
  ggplot(aes(x = sums)) +
  geom_histogram(color = "white")
```
:::


## 4) Box Model

Let $X$ be a discrete random variable with the following _pmf_

$$
f(x) = \begin{cases}
0.1 & \text{for } x = 0 \\
0.2 & \text{for } x = 3 \\
0.3 & \text{for } x = 5 \\
0.4 & \text{for } x = 9
\end{cases}
$$

Write R code to simulate drawing a sample of size 5 with replacement from a box 
associated to the distribution of $X$.

::: {.callout-note collapse="true"}
## Show answers

```{r eval = FALSE}
x <- c(0, 3, 5, 9)
sample(x, size = 5, prob = c(0.1, 0.2, 0.3, 0.4), replace = TRUE)

# equivalent
box <- c(0, 3, 3, 5, 5, 5, 9, 9, 9, 9)
sample(box, size = 5, replace = TRUE)
```
:::


