---
title: "Summarizing Categorical Data"
format: stat20slides-revealjs
chalkboard: true
---

## Agenda

-   Announcements
-   Concept Questions: Handout
-   Worksheet: *Summarizing Categorical Data*
-   Introducing Lab 2.1
-   Work time for Lab 2.1: Understanding Context

## Announcements: Quiz

::: poll
**Quiz-1 next Monday, Sep 8th in class.**

-   Topics:
    -   Understanding the World with Data
    -   Taxonomy of Data
    -   Summarizing Categorical Data (today)
-   Where:
    -   Section 1, 8am in Barker 101
    -   Section 8, 10am in SOCS 60
:::

## Announcements: Quiz

About quizzes:

-   Individual component (25 mins)

-   Group component (15 mins; groups of 3)

-   <u>Handwritten</u> cheat sheet (A4 one-sided)

-   See related posts in Ed

    -   Quiz protocol
    -   Cheat Sheet and Quiz tips

## Announcements (cont'd)

-   WSP-2 due **Tuesday, at 8am** on Gradescope
    -   Taxonomy of Data
    -   Summarizing Categorical Data
    -   *Please scan your paper WS as PDF's*

. . .

-   RQ: Summarizing Numerical Data due **Tuesday at 11:59pm** on Gradescope.

# Practice Concept Questions

Additional handout prepared for Prof. Sanchez's sections (1 & 8)

Answer problems: b, d, f, h, j, l, n, *and* o

This is NOT a worksheet (no need to submit to Gradescope)

```{r}
countdown::countdown(15, top = 0)
```

# Worksheet: *Summarizing Categorical Data*

```{r}
countdown::countdown(25, top = 0)
```

# Break

```{r}
countdown::countdown(5, top = 0)
```

# Introducing Lab 2.1

-   <https://stat20.berkeley.edu/fall-2025/assignments.html>

-   Survey: [tinyurl.com/3hruxwkk](https://tinyurl.com/3hruxwkk)

-   Data glimpse: [tinyurl.com/mzzy7rmc](https://tinyurl.com/mzzy7rmc)

# Lab 1: Understanding the Context of The Data

```{r}
countdown::countdown(25, top = 0)
```

# End of Lecture
